---
title: "p8105_hw1_yl6377"
output: html_document
date: "2026-09-22"
---

## Load the penguins datasets
```{r}
data("penguins", package = "palmerpenguins")
library(ggplot2)
```

## A short description of the penguins dataset
The dataset contains data on penguins from three species: 
`r unique(penguins$species)`. The dataset includes variables such as 
`r names(penguins)`. It has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

## Make a scatterplot
```{r}
ggplot(penguins, aes(x = bill_length_mm,
                     y = flipper_length_mm,
                     color = species)) + geom_point(na.rm=TRUE)

ggsave("penguins_scatterplot.png")
```
## Create a dataframe (Problem 2)
```{r}
library(tidyverse)
set.seed(123)

normal_sample <- rnorm(10)

logical_vector <- normal_sample > 0 

character_vector <- c("A","B","C","D","E","F","G","H","I","J")

factor_vector <- factor(c("low", "medium", "high", "low", "medium", "high", "low", "medium", "high", "low"))

df <- data.frame(
  normal_sample,
  logical_vector,
  character_vector,
  factor_vector
)
```


```{r}
mean(df %>% pull(normal_sample))
mean(df %>% pull(logical_vector))
mean(df %>% pull(character_vector))
mean(df %>% pull(factor_vector))
```
The mean works for the numeric variable. The logical variable also works because R treats TRUE as 1 and FALSE as 0 when calculating a mean. The mean does not work for the character or factor variables because they are not numeric.

```{r, echo=TRUE, results='hide'}
as.numeric(df %>% pull(logical_vector)) 
as.numeric(df %>% pull(character_vector)) 
as.numeric(df %>% pull(factor_vector))
```
The logical vector is converted to 1s and 0s, with TRUE becoming 1 and FALSE becoming 0. The character vector cannot be directly converted to numeric values because its elements are letters. The factor is converted to the underlying numeric codes for its levels rather than the actual factor labels. This helps explain why the mean works for the logical variable but not for the character variable, and why taking the mean of a factor is not appropriate.